---
title: "Quick Mapping Lightning Talk"
author: "Marc Weber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    theme: sandstone
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

## Setup
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(sf)
library(readr)
library(mapview)
library(Rspatialworkshop)
library(ggplot2)
library(leaflet)
library(httr)
```

## Get flat data with coordinates
Use `read_csv` from`readr` package to load a .csv file I keep as part of a spatial data package that you can install with:
```
library(devtools)
install_github("mhweber/Rspatialworkshop")
library(Rspatialworkshop)
```
Then we:

1. Read in the csv
2. Suppress showing column types
3. Select a subset of numerous attributes
4. Filter a small portion of the 2700 gages in the dataset by filtering only gages with drainage areas > 500 sq miles
```{r, warning=FALSE}
gages <- read_csv(system.file("extdata/Gages_flowdata.csv", package = "Rspatialworkshop"),show_col_types = FALSE) %>% dplyr::select(ID=SOURCE_FEA,LON_SITE,LAT_SITE, MeanFlowCFS=AVE, DrainAreaSqMiles=DA_SQ_MILE) %>% 
  dplyr::filter(DrainAreaSqMiles > 500)
glimpse(gages)
```

## Make our data spatial

- We use `st_as_sf` from the `sf` package to promote the data to a `simple feature collection` by:
    + passing the longitude and lattitude fields from our data to use 
    + providing a coordinate reference system - we specify this using an `epsg` code
    + Our `epsg` code is for unprojected NAD83
    + We can find coordinate reference systems (CRS) easily from [spatialreference.org](spatialreference.org)
```{r}
gages <- st_as_sf(gages, coords = c("LON_SITE", "LAT_SITE"), crs = 4269, remove = FALSE)
ggplot() + geom_sf(data=gages)
```

## A simple interactive map using `mapview`
```{r}
mapview(gages)
```

## Customize mapview

- Change the background
- Scale the symbol size by flow rate
```{r}
m <- mapview(gages, map.types = c("Esri.WorldShadedRelief", "OpenStreetMap"), color = "grey40",cex = "MeanFlowCFS")
m
```

## Add Web Map Service (WMS) Layers!
[Web Map Services](https://en.wikipedia.org/wiki/Web_Map_Service) provide pre-rendered map tiles at different scales and are useful as background map layers. Here is my original [Stack Overflow question](https://stackoverflow.com/questions/32960050/loading-wms-layer-in-r-leaflet-using-addwmstiles) from when I first played with adding WMS layers with `mapview` and `leaflet`

### NHD WMS from the National Map
Here we display a background of National Hydrography Dataset stream lines to go with our gages without having to load local files.

This is a super handy way to share your data in an interactive map mashed up with relevant background layers in one file in R, without having to share spatial files.
```{r}
m <- mapview(gages, cex = "MeanFlowCFS")
# Define the NHD WMS service
wms_nhd <- "https://hydro.nationalmap.gov:443/arcgis/services/nhd/MapServer/WmsServer?"
m@map <- m@map %>% 
  addWMSTiles(group = 'USGS HydroCache',
              wms_nhd,layers  = 6,
              options = WMSTileOptions(format = "image/png", transparent = TRUE),
              attribution = "")  %>% mapview:::mapViewLayersControl(names = "NHD")

m
```

### Other WMS Services
We can see other services available on the National Map [here](https://apps.nationalmap.gov/services/) - let's try transportation
```{r}
# Define the WMS service
m <- mapview(gages, cex = "MeanFlowCFS")
wms_trails <- "https://carto.nationalmap.gov:443/arcgis/services/transportation/MapServer/WmsServer?"
m@map <- m@map %>% 
  addWMSTiles(group = 'Trails',
              wms_trails,layers  = 3,
              options = WMSTileOptions(format = "image/png", transparent = TRUE),
              attribution = "")  %>% 
  addWMSTiles(group = 'Trail Labels',
              wms_trails,layers  = 15,
              options = WMSTileOptions(format = "image/png", transparent = TRUE),
              attribution = "") %>%  mapview:::mapViewLayersControl(names = c("Trails","Trail Names"))
                                  
m
```

## Querying and making spatial Web Feature Services (WFS)
Just to show scope of different things we can do linking to REST Services as well - here are links to a couple examples-
[spatial wfs services](https://inbo.github.io/tutorials/tutorials/spatial_wfs_services/) and 
[accessing arcgis rest services](https://community.esri.com/t5/gis-blog/accessing-arcgis-rest-services-using-r/ba-p/898451)

And we can also use REST services for the same WMS layers we displayed in maps above!
See [National Map Services page](https://apps.nationalmap.gov/services/)

We can get a listing of all ESRI REST services [here](https://services.arcgis.com/P3ePLMYs2RVChkJx/ArcGIS/rest/services)

### Query water bodies REST feature
```{r}
url <- parse_url("https://services.arcgis.com/P3ePLMYs2RVChkJx/arcgis/rest/services")
url$path <- paste(url$path, "USA_Water_Bodies/FeatureServer/0/query", sep = "/")
url$query <- list(where = "STATE = 'OR'",
                  outFields = "*",
                  returnGeometry = "true",
                  f = "geojson")
request <- build_url(url)


wb <- read_sf(request)

mapview(gages) + mapview(wb, col.regions='light blue')
```

### Query trails we used earlier
This time use the REST service and a query for 'Pacific Crest National Scenic Trail' - bonus add a default topography brackground with it
```{r}
url <- parse_url("https://carto.nationalmap.gov/arcgis/rest/services")
url$path <- paste(url$path, "transportation/MapServer/34/query", sep = "/")
url$query <- list(where = "Name = 'Pacific Crest National Scenic Trail'",
                  outFields = "*",
                  returnGeometry = "true",
                  f = "json")
request <- build_url(url)

PCT <- read_sf(request)

mapview(PCT, color='dark green',map.types = "OpenTopoMap")
```